---
title: "Natural Populations Preliminary Results"
author: "Daniel Reyes Corral"
date: "7/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("00_packages_v1.R", local = knitr::knit_global())
source("01_rawdata_v1.R", local = knitr::knit_global())
source("02_PCMeans_v1.R", local = knitr::knit_global())
source("03_Variance_Partition_v1.R", local = knitr::knit_global())
```

# Summary results

# Data cleaning
I did a series of PCAs using the variables that were measured continously between 2017 - 2019. I choose these years because 2014 - 2016 were already analysed by Sofia Carvajal. Although we can include these in further analysis (**PTA**) or selection estimates. I removed NAs, check for outliers and corrected some values from the raw data.The clean subset of data was saved as a second table called [**TribulusPC**](https://github.com/Winer-DanielR/Tribulus-Natural-population/tree/master/Data/Processed) this new dataset was used for the following analysis.

## General PCA
The PCA used the following variables:

- *length*
- *width*
- *depth*
- *longest spine*

I did not include *spine tip distance* since I did not measure that trait in 2019 due to the measurements involving a lot of guessing given that most of the mericarps did not have completed spines.

### PCA summary

```{r Summary PCA, echo=FALSE}
summary(Pr)
```

```{r PCA individual, echo=FALSE, fig.cap= "Left: Variance explained per PC. Rigth: Individual PCA as described above."}
par(mfrow=c(1,2))
plot(Pr, type = "l")
biplot(Pr, scale = 0)

```


## PCA means

Then I estimated the means of the PC scores obtained above and plot them in different ways to visualize the patters of variation.

Here I estimated the means per population and years for each island:

```{r PCA means, echo=FALSE}
ggplot(PC_means_year_pop, aes(PC1_mean, PC2_mean)) + geom_point(aes(color = island, size = 5))

```


Here I estimated the means per year and sorted by island. -You can see there is a trend per year that goes in different directions-:

```{r PCA means per year, echo=FALSE}
ggplot(PC_means_year_island, aes(PC1_mean, PC2_mean)) + geom_point(aes(shape = island, size = 5, color = year))
```


## Variance Partition Analysis

I reviewed Sofia's code and use it as a reference to do the partition analysis. I ran a series of individual models per trait. I made sure that population was nested within islands and included island and year as random models. I edited the dataset to make all the other variables factors.

### Descriptive data:

```{r Descriptive, echo=FALSE}
descrip.data
```

```{r Descriptive general, echo=FALSE}
pairs(tribulusPC[,8:14])
```


### Generalized linear model:

lmer(*trait* ~ 1 + (1|year) + (1|island) + (1|island:population), data = tribulusPC, REML = T)

Where *trait* was each of the measured traits:

- [*length*](#summary-results-length)
- [*width*](#summary-results-width)
- [*depth*](#summary-results-depth)
- [*longest spine*](#summary-results-longest-spine)
- [*spine tip distance*](#summary-results-spine-tip-tistance)

For *lower spine* I used a different model since is a binomial trait:

glmer(lower_spine ~ 1 + (1|year) + (1|island) + (1|island:population), data = tribulusPC, na.action = na.omit, family = binomial)

**I did not use *spine position* but I can include that as well**
**I can do a similar analysis per island**

For each trait I used those models I checked assumptions (as Sofia did) and looked at the variance explained by each trait and looked at the significance effect of each random factor: year, island and population within island.

### Summary Results Length:

```{r model summary, echo=FALSE}
summary(lm.length)
```

#### Testing significance of each random effects:
```{r}
ranova(lm.length)
```

#### Checking assumptions *length*:
```{r echo=FALSE}
plot(lm.length)
opar <- par(mfrow=c(2,2))
hist(tribulusPC$length)
plot(fitted(lm.length),resid(lm.length)) 
abline(h=0,lty=2,col="red")
qqnorm(resid(lm.length))
qqline(resid(lm.length), col="red")
hist(resid(lm.length)) 
par(opar)
```

[Top](#generalized-linear-model)

### Summary Results Width
```{r}
summary(lm.width)

```


#### Testing significance of each random effects:
```{r}
ranova(lm.width)

```


#### Checking assumptions *width*:
```{r echo=FALSE}
plot(lm.width)
opar <- par(mfrow=c(2,2))
hist(tribulusPC$width)
plot(fitted(lm.width),resid(lm.width)) 
abline(h=0,lty=2,col="red")
qqnorm(resid(lm.width))
qqline(resid(lm.width), col="red")
hist(resid(lm.width)) 
par(opar)

```

[Top](#generalized-linear-model)


### Summary Results Depth
```{r}
summary(lm.depth)

```


#### Testing significance of each random effects:
```{r}
ranova(lm.depth)

```


#### Checking assumptions *depth*:
```{r echo=FALSE}
plot(lm.depth)
opar <- par(mfrow=c(2,2))
hist(tribulusPC$depth)
plot(fitted(lm.depth),resid(lm.depth)) 
abline(h=0,lty=2,col="red")
qqnorm(resid(lm.depth))
qqline(resid(lm.depth), col="red")
hist(resid(lm.depth)) 
par(opar)

```

[Top](#generalized-linear-model)

### Summary Results Longest Spine
```{r}
summary(lm.longspine)
```


#### Testing significance of each random effects:
```{r}
ranova(lm.longspine)
```


#### Checking assumptions *longest spine*:
```{r echo=FALSE}
plot(lm.longspine)
opar <- par(mfrow=c(2,2))
hist(tribulusPC$longest_spine)
plot(fitted(lm.longspine),resid(lm.longspine)) 
abline(h=0,lty=2,col="red")
qqnorm(resid(lm.longspine))
qqline(resid(lm.longspine), col="red")
hist(resid(lm.longspine)) 
par(opar)
```

[Top](#generalized-linear-model)

### Summary Results Spine Tip Distance

Spine tip distance model removed NAs only uses data from 2017-2018

```{r}
summary(lm.tipdist)
```


#### Testing significance of each random effects:


```{r}
ranova(lm.tipdist)
```



#### Checking assumptions *spine tip distance*:


```{r echo=FALSE}
plot(lm.tipdist)
opar <- par(mfrow=c(2,2))
hist(tribulusPC$spine_tip_distance)
plot(fitted(lm.tipdist),resid(lm.tipdist)) 
abline(h=0,lty=2,col="red")
qqnorm(resid(lm.tipdist))
qqline(resid(lm.tipdist), col="red")
hist(resid(lm.tipdist)) 
par(opar)
```

[Top](#generalized-linear-model)



### Summary Results Lower Spine


```{r}
summary(lm.lowspines)
```


#### Lower Spine plot


```{r}
plot(lm.lowspines)
```

#### Testing significance of each random effects:

This was done manually given that lower spine is a binomial trait. Similar to Sofia's approach I created a model with all the random effects: year, island, island:population and then compared with other models removing each effect and test for significance.


##### Island:Population removed


```{r}
summary(mred)
```

```{r}
anova(mfull, mred)
```


##### Island removed


```{r echo=FALSE}
summary(mred2)
```


```{r echo=FALSE}
anova(mfull, mred2)
```



##### Population removed



```{r}
summary(mred3)
```


```{r}
anova(mfull, mred3)
```


[Top](#generalized-linear-model)

